### 2019-12-27 Statistical Methods/08_Regression.rmd
statistical methods
statistical methods

Vitamin A is required for good health. However, one bite of polar bear liver results in death becasue it contains a high concentration of vitamin A.

Comment on this fact in terms of extrapolation.

@@

Socket et al. (1987) report data related to patterns of residual insulin secretion in children. A portion of the study was concerned with whether age can be used to predict the logarithm of C-peptide concentrations at diagnosis. The observed values are (data file):

Replicate the _LOESS_ smoothed curve in fig 8.4.1.

```{r, echo = T, fig.width=8, fig.height=3.5}
dat <- data.table::fread(paste0(data.dir, "diabetes_sockett_dat.txt"), fill = T)

@@

### 8.16
### 8.17
### 8.18
### 8.19
### 8.20

### 8.21
### 8.22
### 8.23
### 8.24
### 8.25
### 8.26
### 8.27
### 8.28
### 8.29
### 8.30

>>>>

Vitamin A is required for good health. However, one bite of polar bear liver results in death becasue it contains a high concentration of vitamin A. Comment on this fact in terms of extrapolation.

@@

Socket et al. (1987) report data related to patterns of residual insulin secretion in children. A portion of the study was concerned with whether age can be used to predict the logarithm of C-peptide concentrations at diagnosis. The observed values are (data file), Replicate the _LOESS_ smoothed curve in fig 8.4.1.

```{r, echo = T, fig.width=8, fig.height=2.5}
dat <- data.table::fread(paste0(data.dir, "diabetes_sockett_dat.txt"), fill = T)

@@

For the data in the last exercise, use R to verify that a least squares regression line using only X values (age) less than or equal to 7 yields a p-value equal to 0.026 when using the R function _olsch4_. Also verify that the p-value, when using the Theil-Sen estimator, is 0.0233.

```{r, echo = T}
dat2 <- dat[age <= 7]

WRS::olshc4(dat2$age, dat2$pep)

WRS::regci(dat2$age, dat2$pep)
```

### 8.16

For the reading data in Table 8.5, verify that the R function spearci returns a p-value equal to 0.014 and that scorci returns a p-value equal to 0.002. Based on the plot returned by scorci, why is it not suprising that these two function give similar results?

```{r, echo = T, fig.width=8, fig.height=3.5}
dat <- data.table::fread(paste0(data.dir, "read_dat.txt"), fill = T, skip = 13)

spearci( dat[,4], dat[,8])

scorciMC( dat[,4], dat[,8])

### 8.17

Given that $b_1 = -1.5, n = 10, s^2_{y,x} = 35, \sum(X_i - \bar{X})^2 = 140$, assume normality and homosecedasiticy and compute a 0.95 confidence interval for slope, $\beta_1$.

```{r, echo = T}
b1 <- -1.5; n <- 10; se <- sqrt(35 / 140)
alpha <- .05

b1 + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * se
```

### 8.18

Repeat the previous problem, only compute a 0.98 confidence interval.

```{r, echo = T}
b1 <- -1.5; n <- 10; se <- sqrt(35 / 140)
alpha <- 1 - .98

b1 + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * se
```

### 8.19

Based on results convered in the previous chapters, speculate about why the confidence intervals computed in the last two problems might be inaccurate.

_Least squares regression can be negatively impacted by non-normality, heteroscedasticity and outliers._

### 8.20

Assume normality and homoscedasticity and suppose $n = 30, \sum{(X_i)} = 15, \sum(Y_i) = 30, \sum(X_i - \bar{X})(Y_i - \bar{Y}) = 30, \sum(X_i - \bar{X})^2 = 10$

```{r, echo = T}
n <- 30; A <- 30; C <- 10;
xbar <- 15 / 30; ybar <- 30 / 30

slope <- A / C
intercept <- ybar - xbar*slope

Determine the least squares estimates of the slope and intercept.

$\beta_1 = `r slope`, \beta_0 = `r intercept`$

### 8.21

Assume normality and homoscedasticity and suppose $n = 38, \bar{Y} = 20, \sum{X^2_i} = 1,922, \bar{X} = 7, \sum{(X_i - \bar{X})(Y_i - \bar{Y})} = 180, \sum{X_i - \bar{X}^2} = 60, s^2_{X, Y} = 121.$

a.) Determine the least squares estimate of the slope and intercept.

```{r, echo = T}
```

b.) Test the hypothesis $H_0 : \beta_0 = 0, \alpha = 0.05$

```{r, echo = T}
```

c.) Compute a 0.9 confidence interval for $\beta_1$

```{r, echo = T}
```

### 8.22

Assume normality and homoscedasticity and suppose $n = 41, \bar{Y} = 10, \bar{X} = 12, \sum(X_i - \bar{X})(Y_i - \bar{Y}) = 100, \sum(X_i - \bar{X})^2 = 400, s^2_{Y.X} = 144$.

a.) Determine the least squares regression line

b.) Compute a .9 confidence interval for $\beta_1$

### 8.23

Assume normality and homoscedasticity and suppose $n = 18, \beta_1 = 3.1, \sum{(X_i - \bar{X}})^2 = 144, s^2_{X.Y} = 36$.

Compute a 0.95 confidence interval for $\beta_1$.

Would you conclude $\beta_1$ > 2?

### 8.24

Assume normality and homoscedasticity and suppose $n = 20, \beta_0 = 6, \sum{X^2_i} = 169, S^2_{Y.X} = 25, \sum{(X_i - \bar{X}})^2 = 90$.

Compute a .95 confidence interval for $\beta_0$.

### 8.25

Given the following quantities, find the sample correlation coefficent, r, and test $H_0 : \rho = 0$ at the indicated value for $\alpha$.

a.) $n = 27, \sum{(Y_i - \bar{Y})}^2 = 100, \sum{(X_i - \bar{X})^2} = 625, \sum(X_i -\bar{X})(Y_i - \bar{Y})^2 = 200, \alpha = 0.01$

b.) $n = 5, \sum{(Y_i - \bar{Y})}^2 = 16, \sum{(X_i - \bar{X})^2} = 25, \sum(X_i -\bar{X})(Y_i - \bar{Y})^2 = 10, \alpha = 0.05$

### 8.26

### 8.27

### 8.28

### 8.29

### 8.30
